Is the Future of Authenticity In Our Heads?\\\medskip \large Moving Passthoughts From the Lab to the World

### 2017-07-11 paper.tex
Cormac's notes

% Created 2017-07-03 Mon 12:23
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orge209717}

@@

While recent work is heartening, passthoughts remains confined, for now, to the lab.
This paper reviews the immediate questions that must be addressed if passthoughts is to move into everyday life \ref{fig:diagram}.
First, we must understand the real usability properties of passthoughts.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).
Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand 
the space of possible passthoughts,
and how passthoughts change over time (Section 4).

In addition, computing broadly has undergone many sea changes since passthought's initial proposal in 2005.
Biosensing (the sensing of humans) has proliferated throughout our everyday life, due both to the sensors ubiquitous in smartphones, and to general improvements to infrastructures of data collection and machine learning at scale. 
Do these changes make passthoughts less appealing compared to other authentication strategies that these "sea changes" have made possible?

@@

\section{Background}
\label{sec:org6827efb}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgffe92dc}

@@

\subsection{Passthought authentication}
\label{sec:org93d707a}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org09042b7}

@@

\cite{curranpassthoughts} was the first to merge in-ear EEG with passthought authentication,
 using a modified consumer grade EEG device with a single electrode, achieving approximately 80 percent authentication accuracy. 
Ongoing work from the same authors investigates the use of custom-fit earbuds with multiple embedded electrodes \ref{fig:earbud}.
Lending credibility to that study's claim that in-ear EEG could one day become feasible in consumer devices,
United Sciences recently announced a consumer "hearable'' (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics.

\section{User attitudes and perceptions}
\label{sec:orgdbbbb66}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org45e1492}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org6df83c3}

@@

This quote highlights how individuals' life experiences
might shape the way they engage (or refuse to engage) with brain-sensing devices.
In general, this quote and others motivate the need for a rich, qualitative understanding of people's first-hand experiences with brainscanning devices,
as well as data collection,
in order to understand what role BCI applications such as passthoughts could play in day-to-day life.
\section{Diversity and security of passthoughts}
\label{sec:org3330c22}

@@

\section{Two studies on passthoughts}
\label{sec:org7ca404f}

@@

\subsection{A real-time passthought authenticator}
\label{sec:orgf20bd64}

@@

\subsubsection{Study protocol}
\label{sec:orge48e6fe}

@@

\subsubsection{The effect of feedback}
\label{sec:orgf1fddaa}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgef60fad}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orgbea5086}

@@

\subsubsection{Study protocol}
\label{sec:org9be49e3}

@@

\subsubsection{A more diverse corpus}
\label{sec:org4f6869d}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:orgb19bfc8}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:org488cdca}

@@

\subsubsection{Limitations}
\label{sec:org794d9d1}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org12abefb}

@@

\subsection{Privacy}
\label{sec:org13854dd}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:orgac01f42}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:org3cb1742}

@@

\section{Further Future Directions}
\label{sec:orged2d30f}

@@

\subsection{Continuous authentication}
\label{sec:org4bbe60c}

@@

\subsection{Organic passwords}
\label{sec:org4ccf763}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:orgc34f1a2}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org55fcab6}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:orga1d8b9b}

@@

\section{Conclusion}
\label{sec:org74bc9a2}

>>>>

% Created 2017-07-11 Tue 00:34
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orgd1cdd38}

@@

While recent work is heartening, passthoughts remains confined, for now, to the lab.
This paper reviews the immediate questions that must be addressed if passthoughts is to move into everyday life (Figure \ref{fig:diagram}).
First, we must understand the real usability properties of passthoughts.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).
Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand 
the space of possible passthoughts,
and how passthoughts change over time (Section 4).

In addition, computing broadly has undergone many sea changes since passthought's initial proposal in 2005.
Biosensing (the sensing of humans) has proliferated throughout our everyday life, due both to the sensors ubiquitous in smartphones, and to general improvements to infrastructures of data collection and machine learning at scale. 
Is passthoughts more or less appealing in light of other authentication strategies these "sea changes" have made possible?

@@

\section{Background}
\label{sec:orgf24757b}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgc869159}

@@

\subsection{Passthought authentication}
\label{sec:org59cf6e1}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:orgf65c549}

@@

The first attempt to merge in-ear EEG with passthought authentication
used a modified consumer grade EEG device with a single electrode, achieving approximately 80 percent authentication accuracy \cite{curranpassthoughts}.
Ongoing work from the same authors investigates the use of custom-fit earbuds with multiple embedded electrodes \ref{fig:earbud}.
Lending credibility to that study's claim that in-ear EEG could one day become feasible in consumer devices,
United Sciences recently announced a consumer "hearable'' (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics.

\section{User attitudes and perceptions}
\label{sec:orgd0ec90a}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org070ba38}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:orgecc4d41}

@@

This quote highlights how individuals' life experiences
might shape the way they engage (or refuse to engage) with brain-sensing devices.
In general, this quote and others motivate the need for a rich, qualitative understanding of people's first-hand experiences with brainscanning devices,
as well as data collection,
in order to understand what role BCI applications such as passthoughts could play in day-to-day life.
\section{Diversity and security of passthoughts}
\label{sec:org2322423}

@@

\section{Two studies on passthoughts}
\label{sec:orgfeb449a}

@@

\subsection{A real-time passthought authenticator}
\label{sec:org9b0d095}

@@

\subsubsection{Study protocol}
\label{sec:orgbdaa522}

@@

\subsubsection{The effect of feedback}
\label{sec:orge6fa888}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:org1b70139}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:org89ebd65}

@@

\subsubsection{Study protocol}
\label{sec:org51af8ed}

@@

\subsubsection{A more diverse corpus}
\label{sec:orgabae915}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:orgf932e18}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:orge48f8f8}

@@

\subsubsection{Limitations}
\label{sec:org32c6e47}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org8f30435}

@@

\subsection{Privacy}
\label{sec:org792897a}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org9f1d640}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:orge87f2fc}

@@

\section{Further Future Directions}
\label{sec:orgce50b86}

@@

\subsection{Continuous authentication}
\label{sec:orgd749d30}

@@

\subsection{Organic passwords}
\label{sec:orgdc73e84}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:org1391488}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:orgdfab018}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:orgfdf7b6f}

@@

\section{Conclusion}
\label{sec:orga0b4a2b}
